---
title: "scGAD-tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scGAD-tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Recent advancements in single-cell technologies enabled the profiling of 3D genome structures in a single-cell fashion. Quantitative tools are needed to fully leverage the unprecedented resolution of single-cell high-throughput chromatin conformation (scHi-C) data and integrate it with other single-cell data modalities. We developed single-cell gene-body associating domain (scGAD) scores available in R package to explore scHi-C data in units of genes. Proper normalization is investigated to eliminate technical biases.

## Installation

The installation is the same as `BandNorm`, since scGAD is included in this package.

## Format of input

We currently allows two ways input the data:

A path containing all cells in the form of [chr1, binA, chr2, binB, count], the cell looks like below:

```
chr1    0   chr1    10000 9
chr1    10000   chr1    10000 20
chr1    0   chr1    20000 2
chr1    10000 chr1    20000 4
chr1    20000 chr1    20000 22
chr1    10000 chr1    30000 1
chr1    20000 chr1    30000 11
chr1    30000 chr1    30000 197
chr1    10000 chr1    40000 1
chr1    20000 chr1    40000 2
```

Another way is to input a data.frame containing all interactions and the column is like [chr1, binA, binB, count, cell_name]. This way is not recommended, since single-cell Hi-C data is huge. However, it will be fast if your computer can handle the data in this way.

The data should be 10kb resolution. We will allow more formats to be used here, and we will try to allow higher resolution.

For the example dataset, we also have batch information and cell-type information. Note that the cell_name in those external information should be consistent with the hic_df file.

## Use scGAD

We provide a demo scHi-C data named hic_df, sampled 100 cells of Mature Oligodendrocyte, Microglia Etc. and Hippocampal Granule Cell, three cell types and chromosome 1 from [Tan2021](https://www.cell.com/cell/fulltext/S0092-8674(20)31754-2?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS0092867420317542%3Fshowall%3Dtrue) for test run. After obtaining the hic_df file, you can use the `scGAD` function to normalize the data and obtain the scGAD score. The result is similar to scRNAseq, where what you can get is a matrix that rows are genes, and columns are cells.

```{r scGAD}
library(BandNorm)
data(scGADExample)
gad_score = scGAD(hic_df = scgad_df, genes = geneANNO, depthNorm = TRUE)
```

Then, you can use PCA to look at the result, and cells are clearly separated.

```{r PCA Result}
library(ggplot2)
summary = summary[match(colnames(gad_score), summary$cell), ]
gadPCA = prcomp(gad_score)$rotation[, 1:15]
gadPCA = data.frame(gadPCA, cellTypes = summary$`cell-type cluster`)
ggplot(gadPCA, aes(x = PC1, y = PC2, col = cellTypes)) + geom_point() + theme_bw(base_size = 10)
```

